Globally-centered autocovariances in MCMC

### 2020-08-21 Paper/main.tex
Done with round 1. Finished with 31.5 pages. Stopping for today.

Let $F$ be the target distribution with mean $\mu$ defined on the set $\X \subseteq \mathbb{R}^p$ equipped with a countably generated Borel $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}; \; t \in \mathbb{Z}\}$ be the $s^{th}$ $F$-Harris ergodic stationary Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$.
% If $\bar{X} = n^{-1} \sum_{t=1}^{n}X_t$ , then with probability 1, $\bar{X} \to \mu_g$ as $n \to \infty$. \\
The process is covariance stationary so that
the ACvF at lag $k$ depends only on $k$ and is defined as 
%
\[
    \Gamma(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the quality of sample statistics. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator for $\Gamma(k)$ is the sample autocovariance matrix at lag $k$. For the $s$th chain, it is given by
%
\begin{equation} \label{eq:empirical_ACvF}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}} \left(X_{st} - \bar{X}_s \right) \left(X_{s\, t+\abs{k}} - \bar{X}_s \right)^T\,.
\end{equation}
{\color{red} We need to figure out how we want to define autocovariance: with or without absolute value.}
%
% It is a known fact that most of the Markov chains obtained from practical MCMC algorithms are not stationary. Regardless, Equation~\ref{eq:empirical_ACvF} provides the empirical estimator for ACF at a particular lag for the Markov chain with any initial distribution (\cite{geyer2011introduction}).\\

@@

However, there is no unified approach to constructing estimators of $\Gamma(k)$ for multiple-chain implementations. Specifically, parallel MCMC chains are often spread across the state space so as to adequately capture high density areas. For slow mixing Markov chains and multi-modal targets, the chains take time to traverse the space so that estimates of $\mu$ can be vastly different. The sample ACvF from each chain is typically underestimated yielding false sense of security about the quality of process.

@@

In the event that all $m$ Markov chains have been run long enough that $\bar{X}_s \approx \bar{\bar{X}}$ are similar, then  $\hat{\Gamma}_{s} (k) \approx \hat{\Gamma}_{G,s}(k)$. However, for shorter runs or for slow mixing chains, $\Gamma(k)$ is more appropriately estimated by $\hat{\Gamma}_{G,s}$ as it utilizes information from all chains and accounts for disparity between estimates of $\mu$. We quantify the gains in  bias  for the G-ACvF estimator below.
% If the starting points of these parallel Markov chains are sufficiently dispersed, G-ACF provides more realistic estimation of lag covariances. \cite{priestley1981spectral} shows that the empirical estimator in Equation \ref{eq:empirical_ACvF} is biased with the main bias term equal to $\abs{k}\Gamma(k)/n$, ignoring the small order terms arising due to estimation of $\mu$. Though, an unbiased estimator with a divisor of $n - \abs{k}$ instead of $n$ exists, the biased estimator is preferred for its lower mean square error for larger relative $k$  (\cite{priestley1981spectral})
Let 
\[
\Phi^{(q)} = \sum_{ k= -\infty}^{\infty}\abs{k}^q \Gamma(k)\,,
\]
and let $\Phi^{(1)}$ be denoted by $\Phi$.  The proof of the following theorem is available in Appendix~\ref{appendix:bias}.

@@

{\color{orange}MA: I corrected the RHS for the abover equation. Or did you mean to write $\mathbb{E}_F\left[\hat{\Gamma}_{G,s}(k) - \Gamma(k) \right]as LHS?$}

@@

A critical use of autocovariances is in the assessment of Monte Carlo variability of estimators.  Let $g:\X \longrightarrow \mathbb{R}^p$ be an $F$-integrable function so that interest is in estimating
$\mu_g = \mathbb{E}_F[g(X)]$.
% \[
% \mu_g = \mathbb{E}_F[g(X)] = \int_{\X}g(x)F(dx)\;.
% \]
 % 
Set $\{Y_{st}\}_{t \geq 1} = \{g(X_{st})\}_{t \geq 1}$ for all $s = 1, \dots, m$. Let $\bar{Y}_s = n^{-1}\sum_{t=1}^{n}Y_{st}$  and  $\bar{\bar{Y}} = m^{-1}\sum_{s=1}^{m}\bar{Y}_s$. By Birkhoff's ergodic theorem,  $\bar{\bar{Y}} \to \mu_g$ as $n \to \infty$ with probability 1.

An asymptotic sampling distribution may be available via a Markov chain central limit theorem (CLT) if there exists a $p \times p$ positive-definite matrix $\Sigma$ such that
%
\begin{equation}
\label{eq:CLT}
  \sqrt{mn}(\bar{\bar{Y}} - \mu_g) \xrightarrow{d} N(0,\Sigma)\,
\end{equation}
where
\begin{equation}
\label{eq:sigma}
  \Sigma = \sum_{k = -\infty}^{\infty} \Cov_F \left( Y_{11}, Y_{1(1+k)} \right) := \sum_{k = -\infty}^{\infty}\Upsilon (k)
\end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$ in order to assess variability in $\bar{\bar{Y}}$ \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. There is a rich literature on estimating $\Sigma$ for single-chain MCMC implementations. The most common are spectral variance (SV) estimators \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \citep{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from parallel chains. Batch means estimators are computationally more efficient than SV estimators, whereas SV estimators are more reliable \citep{fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV estimator (G-SVE) of $\Sigma$. Using the methods of \cite{heberle2017fast}, we also provide an efficient algorithm for obtaining the G-SV estimator.

@@

\begin{ass}
\label{ass:lag_window}
The lag window $w(x)$ is continuous at $0$ at all but finite number of points. Further, $w(x)$ is a bounded and even function with $w(0)=1, \; {\color{red}\abs{w(x)} < c}$, \; $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W = \sum_{k = -\infty}^{\infty} w(k/b_n)$ is finite.
\end{ass}
% We will consider a range of lag windows $w(x)$; where $w(x)$ is a continuous, even function with $w(0)=1, \abs{w(x)} < c$, $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W_n < \infty$. 
Assumption~\ref{ass:lag_window} is standard \citep[see]{ande:1971} and is satisfied by most lag windows. We will use the following popular Bartlett lag window in our simulations:
\[
w(x) = \begin{cases}
1 - \abs{x} & \textrm{for } \abs{x} \leq 1 \\
0 & \textrm{otherwise}\,.
\end{cases}
\]

@@

The simulation terminates when the ESS is greater than a pre-specified, theoretically motivated, lower-bound $W_{p}$. 
% For a multi-modal target distribution, the single chain estimators for $\Sigma$ underestimate the true generalized variance; giving an unrealistic assurance of high ESS for a given $n$. 
%
In our setting of $m$ parallel chains of $n$ samples each, define
%
\[
\hat{\Lambda}_{mn} = \dfrac{1}{m(n-1)}\sum_{s=1}^{m}\sum_{t=1}^{n}(Y_{st} - \bar{Y}_s)(Y_{st} - \bar{Y}_s)^T\, ,
\]
and estimate ESS with
\[
\widehat{\textrm{ESS}}_G = mn\left(\dfrac{|\hat{\Upsilon}(0)|}{|\hat{\Sigma}_{G}|}\right)^{1/p}\, .
\]
We use the locally-centered $\hat{\Upsilon}(0)$ to estimate $\Upsilon(0)$ instead of $\hat{\Upsilon}_G(0)$ when calculating ESS. Both $\hat{\Upsilon}(0)$ and $\hat{\Upsilon}_G(0)$ are consistent  for $\Upsilon(0)$; however using $\hat{\Upsilon}(0)$ produces lower sample sizes when the parallel chains do not agree, indicating that more MCMC draws are needed.

@@

The quality of estimation of $\Sigma$ and $\text{ESS}$ is assessed by running plots from 50 replications of run length 50000. 
 In Figure~\ref{fig:var-frob_n_ess}, we present running plots of $\log(\|\hat{\Sigma}\|_F)$ for both $\hat{\Sigma}_G$ and $\hat{\Sigma}_A$ and the running plots of $\log(\widehat{\text{ESS}})$ for both  $\widehat{\text{ESS}}_G$ and $\widehat{\text{ESS}}_A$. It is evident that the locally centered estimator of $\Sigma$ severely underestimates the truth, leading to an overestimation of ESS. The GSV estimator is able to estimate $\Sigma$ more accurately early on, safeguarding against early termination using ESS. 
\begin{figure}[htbp]
    \centering
      \includegraphics[width = 2.2in]{plots/var-frob.pdf} 
      \includegraphics[width = 2.2in]{plots/var-ess.pdf}         
    \caption{VAR(1). (Left) Running plot for logarithm of Frobenius norm of A-SV and G-SV.(Right) Running plot for logarithm of $\widehat{\textrm{ESS}}/mn$ using A-SV and G-SV.}
    \label{fig:var-frob_n_ess}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width = .8\textwidth]{plots/boom-all_1_3_8.pdf}\\
    \includegraphics[width = .8\textwidth]{plots/boom-all_1_10_7.pdf}    
    % \includegraphics[width = 0.6\textwidth ]{plots/boom-acf_1_10_7_n1000.pdf}

@@

\begin{table}[h]
\parbox{.45\linewidth}{
\centering
\small
\begin{tabular}{|c|c|c|c|c|}
\hline
 $n$ & \multicolumn{2}{|c|}{$m = 2$} & \multicolumn{2}{|c|}{$m=5$}\\
 \hline
 & A-SV & G-SV & A-SV & G-SV \\
 \hline
 5000 & 0.595 &  0.700 & 0.402 &  0.640\\
 10000 & 0.563 &  0.665 & 0.59 &  0.739\\
 50000 & 0.775 &  0.814 & 0.807 &  0.864\\
 100000 & 0.847 &  0.864 & 0.884 &  0.902\\
\hline
\end{tabular}
\caption{Setting-1}
\label{tab:boom-coverage_1}
}
\hfill
\parbox{.45\linewidth}{
\centering
\small
\begin{tabular}{|c|c|c|c|c|}
 \hline
 $n$ & \multicolumn{2}{|c|}{$m = 2$} & \multicolumn{2}{|c|}{$m=5$}\\
 \hline
 & A-SV & G-SV & A-SV & G-SV \\
 \hline
 1000 &  0.856 &  0.868 & 0.895 &  0.910\\
 5000 & 0.921 &  0.925 & 0.910 &  0.915\\
 10000 & 0.928 &  0.93 & 0.919 &  0.926\\
 50000 & 0.943 &  0.944 & 0.951 &  0.952\\
\hline
\end{tabular}
\caption{Setting-2}
\label{tab:boom-coverage_2}
}
\end{table}

@@

\begin{figure}[h]
    \centering
      \includegraphics[width = .35\textwidth]{plots/boom-ess_1_3_8_m5.pdf}
      \includegraphics[width = .35\textwidth]{plots/boom-ess_1_10_7_m5.pdf}

@@

For our third example, we consider a real-life example of sensor locations previously discussed by \cite{ihler2005nonparametric}. The goal is to identify unknown sensor locations using noisy distance data. This problem is specifically interesting in our case because the marginal posterior distribution for missing sensor locations is multi-modal for all locations.

Following \cite{tak2018repelling}, we assume that there are six sensors scattered on a planar region where $x_i = (x_{i1}, x_{i2})^T$ denote the $2d$ coordinates of $i^{th}$ sensor. Let $y_{ij} = (y_{ji})$ denote the distance between the sensors $x_i$ and $x_j$. The distance between $x_i$ and $x_j$ is observed with probability $\pi (x_i, x_j) = \exp{\|x_i - x_j\|^2 / 2R^2}$ and with a Gaussian measurement error of $\sigma^2$. Let $z_{ij}$ denote the indicator variable which is equal to 1 when the distance between $x_i$ and $x_j$ is observed. The probability model is then,
%
\begin{align*}
    z_{ij} \mid x_1, ..., x_6 & \sim \text{Bernoulli}\left(\exp\left(\dfrac{-\|x_i - x_j\|^2}{2R^2}\right)\right)\\
    y_{ij} \mid w_{ij} = 1, x_1, ..., x_6 &\sim N(\|x_i - x_j\|^2, \sigma^2)
\end{align*}
%
\cite{ahn2013distributed} suggested the value of $R = 0.3$ and $\sigma = 0.02$. We use a Gaussian prior for the unknown locations with  mean equal to $(0,0)$ and covariance matrix equal to $100 I_2$. $y_{ij}$ is specified only if $w_{ij} = 1$. We follow the Markov chain structure as described by \cite{tak2018repelling} and sample from the four bivariate conditionals for each sensor location using a Gibbs sampler. In their paper on Repelling Attractive Metropolis (RAM) algorithms, \cite{tak2018repelling} compare the performance of different sampling techniques and show that RAM improves the acceptance rate by a factor of at least 5.5 over Metropolis using the same jumping scale. RAM algorithm supplies Markov chains with higher jumping frequency between the modes.

\begin{figure}[h]
    \centering
    \includegraphics[width = .32\textwidth]{plots/sensor-trace_loc1.pdf}
      \includegraphics[width = .32\textwidth]{plots/sensor-acf_n5e3.pdf}
      \includegraphics[width = .32\textwidth]{plots/sensor-acf_n5e4.pdf}    
    \caption{Sensor: Trace plots for $x$ (left) and $y$ (right) coordinate of sensor location-1.}
    \label{fig:sensor-trace}
\end{figure}

This is an 8-dimensional estimation problem where the components corresponds to x and y coordinates of four unknown sensor locations. We will use the RAM algorithm with a jumping scale equal to $0.5$ to sample five parallel Markov chains with well-separated starting points. The total simulation size for each chain is fixed at $100,000$. Since the truth about actual mean and asymptotic variance is not known in this case, we are not interested in coverage probabilities.

To illustrate the \textit{sticky} nature of the Markov chains, we plot the evolution of two chains (starting near different modes) with time. Figure~\ref{fig:sensor-trace} shows the trace plot of x and y coordinate of sensor location-1. It is evident that the marginal posterior for both the components is bimodal. Similarly, the other three locations also have a bimodal marginal distribution in each component.  Observe that each chain explores one particular mode for a long time before jumping to another mode.

% \begin{figure}[h]
%     \centering
%       \includegraphics[width = .33\textwidth]{plots/sensor-acf_n5e3.pdf}
%       \includegraphics[width = .33\textwidth]{plots/sensor-acf_n5e4.pdf}

%     \caption{Sensor: ACF (orange) and G-ACF (blue) for individual chains (dotted lines) and average over all chains (solid line) for $n = 5000$ (left) and $n = 50000$ (right).}
%     \label{fig:sensor-acf}
% \end{figure}

@@

Figure~\ref{fig:sensor-acf} plots the locally and globally centered ACF estimators for individual chains  as well as the average over all the chains. The thick solid line plots the averaged estimator and the thin dotted lines correspond to ACF calculated using single chains. Figure~\ref{fig:sensor-frob_n_ess} presents the running plot of $\log \|{\Sigma}\|_F$ and $\log \textrm{ESS}/mn$ estimated using A-SVE and G-SVE along with standard errors from $10$ replications for each value of $n$. In both the plots, G-SVE and $\widehat{\textrm{ESS}}_G/mn$ reach the convergence value faster.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width = 0.7\textwidth]{plots/poisson-trace_n50000.pdf}
    \caption{Trace plot for second (left) and third (right) component of two chains started from random points.The two colors denote the two different chains.}
    \label{fig:poisson-trace}
\end{figure}

@@

\begin{figure}[h]
    \centering
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n1000.pdf}    
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n10000.pdf}      
    \caption{Change point: ACF (left) and G-ACF (right) for individual chains and average over $m$ chains for $n = 1000$ (left) and $n = 10000$ (right).}
    \label{fig:poisson-acf}
\end{figure}

@@

\begin{figure}
    \centering
      \includegraphics[width = .6\textwidth]{plots/magnolia-acf_n100.pdf}
      \includegraphics[width = .6\textwidth]{plots/magnolia-acf_n1000.pdf}

>>>>

Let $F$ be the target distribution with mean $\mu$ defined on the set $\X \subseteq \mathbb{R}^p$ equipped with a countably generated Borel $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}; \; t \in \mathbb{Z}\}$ be the $s^{th}$ $F$-Harris ergodic stationary Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$.
% If $\bar{X} = n^{-1} \sum_{t=1}^{n}X_t$ , then with probability 1, $\bar{X} \to \mu_g$ as $n \to \infty$. \\
The process is covariance stationary so that
the ACvF at lag $k$ depends only on $k$ and is defined as 
%
\[
    \Gamma(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the quality of sample statistics. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator for $\Gamma(k)$ is the sample autocovariance matrix at lag $k$. For the $s$th chain, it is given by
%
\begin{equation} \label{eq:empirical_ACvF}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}} \left(X_{st} - \bar{X}_s \right) \left(X_{s\, t+\abs{k}} - \bar{X}_s \right)^T\,.
\end{equation}

%
% It is a known fact that most of the Markov chains obtained from practical MCMC algorithms are not stationary. Regardless, Equation~\ref{eq:empirical_ACvF} provides the empirical estimator for ACF at a particular lag for the Markov chain with any initial distribution (\cite{geyer2011introduction}).\\

@@

However, there is no unified approach to constructing estimators of $\Gamma(k)$ for multiple-chain implementations. Specifically, parallel MCMC chains are often spread across the state space so as to adequately capture high density areas. For slow mixing chains and multi-modal targets, the chains take time to traverse the space so that estimates of $\mu$ are vastly different. The ACvF from each chain is typically underestimated yielding false sense of security about the quality of process.

@@

In the event that all $m$ Markov chains have been run long enough that $\bar{X}_s \approx \bar{\bar{X}}$ are similar, then  $\hat{\Gamma}_{s} (k) \approx \hat{\Gamma}_{G,s}(k)$. However, for shorter runs or for slow mixing chains, $\Gamma(k)$ is more appropriately estimated by $\hat{\Gamma}_{G,s}$ as it utilizes information from all chains and accounts for disparity between estimates of $\mu$. We quantify the gains in  bias  for the G-ACvF estimator below. Let 
\[
\Phi^{(q)} = \sum_{ k= -\infty}^{\infty}\abs{k}^q \Gamma(k)\,,
\]
and let $\Phi^{(1)}$ be denoted by $\Phi$.  The proof of the following theorem is available in Appendix~\ref{appendix:bias}.

@@

@@

A critical use of autocovariances is in the assessment of Monte Carlo variability of estimators.  Let $g:\X \longrightarrow \mathbb{R}^p$ be an $F$-integrable function so that interest is in estimating
$\mu_g = \mathbb{E}_F[g(X)]$.
% \[
% \mu_g = \mathbb{E}_F[g(X)] = \int_{\X}g(x)F(dx)\;.
% \]
 % 
Set $\{Y_{st}\}_{t \geq 1} = \{g(X_{st})\}_{t \geq 1}$ for all $s = 1, \dots, m$. Let $\bar{Y}_s = n^{-1}\sum_{t=1}^{n}Y_{st}$  and  $\bar{\bar{Y}} = m^{-1}\sum_{s=1}^{m}\bar{Y}_s$. By Birkhoff's ergodic theorem,  $\bar{\bar{Y}} \to \mu_g$ as $n \to \infty$ with probability 1. An asymptotic sampling distribution may be available via a Markov chain central limit theorem (CLT) if there exists a $p \times p$ positive-definite matrix $\Sigma$ such that
%
\begin{equation}
\label{eq:CLT}
  \sqrt{mn}(\bar{\bar{Y}} - \mu_g) \xrightarrow{d} N(0,\Sigma)\, \quad \text{ where, }
\end{equation}
% where
\begin{equation}
\label{eq:sigma}
  \Sigma = \sum_{k = -\infty}^{\infty} \Cov_F \left( Y_{11}, Y_{1(1+k)} \right) := \sum_{k = -\infty}^{\infty}\Upsilon (k)
\end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$ in order to assess variability in $\bar{\bar{Y}}$ \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. There is a rich literature on estimating $\Sigma$ for single-chain MCMC implementations. The most common are spectral variance (SV) estimators \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \citep{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from parallel chains. Batch means estimators are computationally more efficient than SV estimators, whereas SV estimators are more reliable \citep{fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV estimator (G-SVE) of $\Sigma$. Using the methods of \cite{heberle2017fast}, we also provide an efficient algorithm for obtaining the G-SV estimator.

@@

\begin{ass}
\label{ass:lag_window}
The lag window $w(x)$ is continuous at $0$ at all but finite number of points, is a bounded and even function with $w(0)=1$, \; $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $\sum_{k = -\infty}^{\infty} w(k/b_n) < \infty$.
\end{ass}

Assumption~\ref{ass:lag_window} is standard \citep[see]{ande:1971} and is satisfied by most lag windows. We will use the following popular Bartlett lag window in our simulations:
\[
w(x) = \begin{cases}
1 - \abs{x} & \textrm{for } \abs{x} \leq 1 \\
0 & \textrm{otherwise}\,.
\end{cases}
\]

@@

The simulation terminates when the ESS is greater than a pre-specified, theoretically motivated, lower-bound $W_{p}$. 
% For a multi-modal target distribution, the single chain estimators for $\Sigma$ underestimate the true generalized variance; giving an unrealistic assurance of high ESS for a given $n$. 
%
In our setting of $m$ parallel chains of $n$ samples each, we
%
% \[
% \hat{\Lambda}_{mn} = \dfrac{1}{m(n-1)}\sum_{s=1}^{m}\sum_{t=1}^{n}(Y_{st} - \bar{Y}_s)(Y_{st} - \bar{Y}_s)^T\, ,
% \]
 estimate ESS with
\[
\widehat{\textrm{ESS}}_G = mn\left(\dfrac{|\hat{\Upsilon}(0)|}{|\hat{\Sigma}_{G}|}\right)^{1/p}\, .
\]
We use the locally-centered $\hat{\Upsilon}(0)$ to estimate $\Upsilon(0)$ instead of $\hat{\Upsilon}_G(0)$ when calculating ESS. Both $\hat{\Upsilon}(0)$ and $\hat{\Upsilon}_G(0)$ are consistent  for $\Upsilon(0)$; however using $\hat{\Upsilon}(0)$ produces lower sample sizes when the parallel chains do not agree, indicating that more MCMC draws are needed.

@@

The quality of estimation of $\Sigma$ and $\text{ESS}$ is assessed by running plots from 50 replications of run length 50000. 
 In Figure~\ref{fig:var-frob_n_ess}, we present running plots of $\log(\|\hat{\Sigma}\|_F)$ for both $\hat{\Sigma}_G$ and $\hat{\Sigma}_A$ and the running plots of $\log(\widehat{\text{ESS}})$ for both  $\widehat{\text{ESS}}_G$ and $\widehat{\text{ESS}}_A$. It is evident that the locally centered estimator of $\Sigma$ severely underestimates the truth, leading to an overestimation of ESS. The GSV estimator is able to estimate $\Sigma$ more accurately early on, safeguarding against early termination using ESS. 
\begin{figure}[htbp]
    \centering
      \includegraphics[width = 2.0in]{plots/var-frob.pdf} 
      \includegraphics[width = 2.0in]{plots/var-ess.pdf}         
    \caption{VAR(1). (Left) Running plot for logarithm of Frobenius norm of A-SV and G-SV.(Right) Running plot for logarithm of $\widehat{\textrm{ESS}}/mn$ using A-SV and G-SV.}
    \label{fig:var-frob_n_ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .75\textwidth]{plots/boom-all_1_3_8.pdf}\\
    \includegraphics[width = .75\textwidth]{plots/boom-all_1_10_7.pdf}    
    % \includegraphics[width = 0.6\textwidth ]{plots/boom-acf_1_10_7_n1000.pdf}

@@

\begin{table}[htbp]
\parbox{.45\linewidth}{
\centering
\small
\begin{tabular}{|c|c|c|c|c|}
\hline
 $n$ & \multicolumn{2}{|c|}{$m = 2$} & \multicolumn{2}{|c|}{$m=5$}\\
 \hline
 & A-SV & G-SV & A-SV & G-SV \\
 \hline
 5000 & 0.595 &  0.700 & 0.402 &  0.640\\
 10000 & 0.563 &  0.665 & 0.59 &  0.739\\
 50000 & 0.775 &  0.814 & 0.807 &  0.864\\
 100000 & 0.847 &  0.864 & 0.884 &  0.902\\
\hline
\end{tabular}
\caption{Setting-1}
\label{tab:boom-coverage_1}
}
\hfill
\parbox{.45\linewidth}{
\centering
\small
\begin{tabular}{|c|c|c|c|c|}
 \hline
 $n$ & \multicolumn{2}{|c|}{$m = 2$} & \multicolumn{2}{|c|}{$m=5$}\\
 \hline
 & A-SV & G-SV & A-SV & G-SV \\
 \hline
 1000 &  0.856 &  0.868 & 0.895 &  0.910\\
 5000 & 0.921 &  0.925 & 0.910 &  0.915\\
 10000 & 0.928 &  0.93 & 0.919 &  0.926\\
 50000 & 0.943 &  0.944 & 0.951 &  0.952\\
\hline
\end{tabular}
\caption{Setting-2}
\label{tab:boom-coverage_2}
}
\end{table}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .30\textwidth]{plots/boom-ess_1_3_8_m5.pdf}
      \includegraphics[width = .30\textwidth]{plots/boom-ess_1_10_7_m5.pdf}

@@

Consider the sensor network localization problem discussed in \cite{ihler2005nonparametric}. The goal is to identify unknown sensor locations using noisy distance data. We use the data
% This problem is specifically interesting in our case because the marginal posterior distribution for missing sensor locations is multi-modal for all locations. 
%
We use the data and the model as specified by \cite{tak2018repelling}. There are four sensors scattered on a planar region where $x_i = (x_{i1}, x_{i2})^T$ denotes the coordinates of the $i${th} sensor. Let $y_{ij} = (y_{ij})$ denotes the distance between the sensors $x_i$ and $x_j$ and if observed yields $z_{ij} = 1$, otherwise, $z_{ij} = 0$. The complete model is then,
\begin{align*}
    z_{ij} \mid x_1, ..., x_4 & \sim \text{Bernoulli}\left(\exp\left(\dfrac{-\|x_i - x_j\|^2}{2R^2}\right)\right)\\
    y_{ij} \mid z_{ij} = 1, x_i,x_j &\sim N(\|x_i - x_j\|^2, \sigma^2)
\end{align*}
%
\cite{tak2018repelling} set $R = 0.3$ and $\sigma = 0.02$ and use a Gaussian prior for the unknown locations with  mean equal to $(0,0)$ and covariance matrix equal to $100 I_2$. $y_{ij}$ is specified only if $w_{ij} = 1$. The resulting $8$-dimensional posterior of $(x_1, \dots, x_4)$ is intractable with unknown full conditionals. A Metropolis-within-Gibbs type sampler is implemented with each full conditional employing the repelling attractive Metropolis (RAM) algorithm of \cite{tak2018repelling}. The RAM algorithm supplies Markov chains with higher jumping frequency between the modes.

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .32\textwidth]{plots/sensor-trace_loc1.pdf}
      \includegraphics[width = .32\textwidth]{plots/sensor-acf_n5e3.pdf}
      \includegraphics[width = .32\textwidth]{plots/sensor-acf_n5e4.pdf}    
    \caption{Sensor: Trace plot of $x_{11}$ (left). Locally-centered ACF (solid orange) and G-ACF (solid blue) at $n = 5000$ (middle) and $n = 50000$ (right). Dashed lines are single-chain estimated ACFs.}
    \label{fig:sensor-trace}
\end{figure}

@@

We run $m = 5$ parallel Markov chains with well-separated starting points. The total simulation size for each chain is fixed at $10^5$. Coverage probabilities are not estimable since the true posterior mean is unknown. Trace plot of $x_{11}$ for two Markov chains is shown in Figure~\ref{fig:sensor-trace}. Figure~\ref{fig:sensor-trace} also  plots locally and globally centered ACFs. As seen in previous examples, early estimates of locally-centered ACFs are must smaller than later estimates. The globally-centered ACFs, on the other hand are similar at small and large sample sizes. Figure~\ref{fig:sensor-frob_n_ess} presents the running plot of $\log \|{\Sigma}\|_F$ and $\log \textrm{ESS}/mn$ using A-SV and G-SV  estimators along with standard errors from $10$ replications. In both the plots, G-SV and $\widehat{\textrm{ESS}}_G/mn$ reaches stability significantly earlier than A-SV.

@@

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.7\textwidth]{plots/poisson-trace_n50000.pdf}
    \caption{Trace plot for second (left) and third (right) component of two chains started from random points.The two colors denote the two different chains.}
    \label{fig:poisson-trace}
\end{figure}
\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n1000.pdf}  \\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n10000.pdf}      
    \caption{Change point: ACF (left) and G-ACF (right) for individual chains and average over $m$ chains for $n = 1000$ (left) and $n = 10000$ (right).}
    \label{fig:poisson-acf}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/magnolia-acf_n100.pdf}\\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/magnolia-acf_n1000.pdf}
